---
title: "Reproducible_Research_P_1"
author: "Saad"
date: "1/28/2022"
output:
  html_document: 
   keep_md: TRUE
---

## Loading and preprocessing the data

Loading relevant dependencies

```{r echo=FALSE, message=FALSE}
library(lubridate)
library(dplyr)
library(tidyverse)
library(mice)
library(ggplot2)
```

Loading Data

```{r reading, include=TRUE, cache=TRUE}
watch_ds <- read.csv("activity.csv")
summary(watch_ds)
```

Processing Data

```{r date, cache=TRUE} 
watch_ds$date <- ymd(watch_ds$date)
watch_ds$day <- wday(watch_ds$date, label = T)
```

## What is mean total number of steps taken per day?

Carving out the Non-NA values  
The total number of steps taken per day

```{r separation, cache=TRUE}
Day_average <- watch_ds %>% 
        group_by(date) %>% 
        summarise(total_steps = sum(steps, na.rm = T), na = mean(is.na(steps)))
Day_average
Non_NA_Average <- aggregate(watch_ds$steps ~ watch_ds$date, FUN = sum, )
colnames(Non_NA_Average) <- c("DATE", "T_Steps")
Non_NA_Average
```


## Data Visualization

Grphically seeing the data

```{r histogram, fig.width= 12, cache=TRUE}
hist(Non_NA_Average$T_Steps, col = "yellow", labels = T, 
     xlab = "Steps", ylab = "Day(Without NA's)" , 
     main = "Total Steps Taken Per Day", 
     ylim = c(0,40))
```

Mean & Median of total No of Steps Taken per Day including NA's

```{r summary, cache=TRUE}
mean_t_steps <- mean(Non_NA_Average$T_Steps)
mean_t_steps <- as.integer(mean_t_steps)
mean_t_steps
median_t_steps <- as.integer(median(Non_NA_Average$T_Steps))
median_t_steps
```
The average number of steps taken each day was **`r mean_t_steps`** steps
The median number of steps taken each day was **`r median_t_steps`** steps


## What is the average daily activity pattern?

Devising the Average Daily Step Pattern

```{r average_day, fig.width=12, cache=TRUE}
A_Mean_Day <- watch_ds %>% 
        group_by(interval) %>% 
        summarise(routine = mean(steps, na.rm = T))

plot(x = 1:nrow(A_Mean_Day), y = A_Mean_Day$routine, type = "l", 
     col = "black", xaxt = "n", xlab = "Intervals", 
     ylab = "Average of the days in that interval", 
     ylim = c(0,210))
axis(1, labels = A_Mean_Day$interval[seq(1, 288, 12)], 
     at = seq_along(A_Mean_Day$interval)[seq(1,288,12)])

```

Which 5-minute interval contains the maximum number of steps?

```{r maximum 5_min interval, cache=TRUE}
max_point <- filter(A_Mean_Day, routine == max(A_Mean_Day$routine))
max_point
```

Interval **`r max_point$interval`** contains an average maximum number of 
steps of about **`r round(max_point$routine,2)`**


# Working on Imputing the Data

```{r NA_identifying, }
activity <- read.csv("activity.csv")
activity$steps[which(is.na(activity$steps))] = mean(activity$steps, na.rm = T)
``` 

```{r Average_imputed_day,cache=TRUE}
Average_Day_impute <- activity %>% 
        group_by(interval) %>% 
        summarise(routine = mean(steps))
```

```{r imputed_cleaning, }
activity$date <- ymd(activity$date)
activity$day <- wday(activity$date, label = T)
imput_Average <- aggregate(activity$steps ~ activity$date, FUN = sum, )
colnames(imput_Average) <- c("DATE", "T_Steps")
```

```{r visualization}
plot(x = 1:nrow(Average_Day_impute), y = Average_Day_impute$routine, type = "l", 
     col = "blue", xaxt = "n", xlab = "Intervals", 
     ylab = "Average of the days in that interval", 
     ylim = c(0,210))
axis(1, labels = Average_Day_impute$interval[seq(1, 288, 12)], 
     at = seq_along(Average_Day_impute$interval)[seq(1,288,12)])
hist(imput_Average$T_Steps, col = "yellow", labels = T, 
     xlab = "Steps", ylab = "Day(Without NA's)" , 
     main = "Total Steps Taken Per Day", 
     ylim = c(0,40))
```
Mean & Median of total No of Steps After Imputation

```{r summary_imput, cache=F}
mean_t_imp_steps <- mean(imput_Average$T_Steps)
mean_t_steps <- as.numeric(mean_t_imp_steps)
mean_t_imp_steps
median_t_imp_steps <- as.integer(median(imput_Average$T_Steps))
median_t_imp_steps
```
The average number of steps taken each day was **`r mean_t_imp_steps`** steps
The median number of steps taken each day was **`r median_t_imp_steps`** steps

```{r Day_Clasifying, }
activity$Day_class <- ifelse(activity$day == "Sun" | activity$day == "Sat", 
                             "off_day", ifelse(activity$day == "Fri", 
                                               "Half_day", "Work_day"))
activity$Day_class <- as.factor(activity$Day_class)
activity$steps <- as.numeric(activity$steps)
```

```{r Day_Classified_Mean}
Day_class_Mean <- activity %>% 
        group_by(interval, Day_class) %>% 
        summarise(routine = mean(steps))
```

```{r Panel_Plot, }
qplot(interval, routine, data = Day_class_Mean,geom = "line" ,
      facets = Day_class ~.)
```


